---
title: "Lab 13"
author: "Nick Gulotta"
date: "12/7/2021"
output: github_document
html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r data, include=T, message=F, warning=F}
require(marked)
require(readr)
require(gridExtra)
require(dplyr)

#data
cjs.data <- read.table("/Users/nickg/OneDrive/Desktop/R projects/Population-dynamics-FANR/data/cawa-cjs.inp", sep=" ",
                               colClasses=c("character","character"),
                               col.names=c("ch", "count"))

js.data <- read.table("/Users/nickg/OneDrive/Desktop/R projects/Population-dynamics-FANR/data/CH-SO-Dean-AllYears.inp", sep=" ",
                               colClasses=c("character","character"),
                               col.names=c("ch", "count"))
```


### **Exercise I**

#### **$Φ(.)p(.)$ – No variation in apparent survival**

```{r, mod1, include=T, warning=F,message=F}
phi0.p0 <- crm(data=cjs.data, model="CJS", hessian=TRUE,
model.parameters=list(
Phi=list(formula=~1), ## No variation
p=list(formula=~1))) ## No variation

phi0.p0

predict(phi0.p0)
```

#### **$Φ(t)p(.)$ – Temporal variation in apparent survival**

```{r, mod2, include=TRUE, warning=F, message=F}
phiTime.p0 <- crm(data=cjs.data, model="cjs", hessian=TRUE,
model.parameters=list(
Phi=list(formula=~time), ## Temporal variation
p=list(formula=~1))) ## No variation

phiTime.p0

predict(phiTime.p0)
```

#### **$Φ(.)p(t)$ – Temporal variation in capture probability**

```{r, mod3, include=T, warning=F, message=F}
pTime.p0<-crm(data=cjs.data, model="cjs", hessian=TRUE,
model.parameters=list(
Phi=list(formula=~1), ## no variation
p=list(formula=~time))) ## variation

pTime.p0

predict(pTime.p0)
```
#### **$Φ(t)p(t)$ – Temporal variation in apparent survival and capture probability**

```{r, mod4, include=T, warning=F, message=F}
pPhiTime.p0<-crm(data=cjs.data, model="cjs", hessian=TRUE,
model.parameters=list(
Phi=list(formula=~time), ## variation
p=list(formula=~time))) ## variation

pPhiTime.p0

pPhiTime.p<-predict(pPhiTime.p0)
```

#### **Which model is the best?**

**The first model that had no variation was the best model since it had the lowest AIC score.**

```{r table1, include=T, warning=F, message=F, echo=FALSE}
# create HTML table using kableExtra
library(kableExtra)
# create HTML table using kableExtra
library(kableExtra)
require(knitr)

#Extract estimates from model w/no variation
pred.phi0.p0<-predict(phi0.p0)
pred.phi0<-pred.phi0.p0[["Phi"]]
pred.p0<-pred.phi0.p0[["p"]]

#make tibbles to join
pred.phi0<-as_tibble(pred.phi0)
pred.p0<-as_tibble(pred.p0)

#full join both datasets for table
d2<-full_join(pred.phi0,pred.p0)

Model<-c("No variation in apprent surival or capture probability"," ")
Term<-c("Phi","p")
AIC<-c(168.64," ")


tab<-data.frame(Model,Term,d2,AIC)
colnames(tab) <- c("Model","Term" ,"occ", "Estimate", "SE", "lcl", "ucl", "AIC") ## Rename


#table 
options(knitr.kable.NA = "") # leave NA cells empty
knitr::kable(tab, digits = 2,booktabs=T, align="c",
       caption = "<center><strong>Table 1. Results for distance sampling of gazelle for half-normal and half-hazard models.</strong></center>")%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"))
```

### **Exercise II**

#### **Jolly-Serber w/constant capture probability (p), constant entrance probabilities (pent, $b_t$), and temporal variation in apparent survival**

```{r, mod5, include=T, warning=F, message=F}
js.phi0.pent0.p0 <- crm(data=js.data, model="JS", hessian=TRUE,
                        model.parameters=list(
                          Phi=list(formula=~time), ## No variation
                          pent=list(formula=~1), ## No variation
                          p=list(formula=~1))) ## No variation

js0est <- predict(js.phi0.pent0.p0)
js0est


n <- nrow(js.data) # number of individuals captured
Nsuper <- n+js0est$N$estimate # Super-population size
Nsuper

b <- js0est$pent$estimate # Entrance probabilities after first time period
b0 <- 1-sum(b) # Compute first entrance probability
round(b, digits=16)

R <- Nsuper*b # Recruits
R


Phi <- js0est$Phi$estimate ## Apparent survival probability
Phi

```

#### **Interpretation of estimates**
**The capture probability was p= 0.26, and the estimated number of individuals not detected was n= 72.82. **

#### **Compute: super population size, number of recruits in each time interval, abundance at each time point**

```{r, compute, include=T}
n <- nrow(js.data) # number of individuals captured
Nsuper <- n+js0est$N$estimate # Super-population size
Nsuper

b <- js0est$pent$estimate # Entrance probabilities after first time period
b0 <- 1-sum(b) # Compute first entrance probability
round(b, digits=16)

R <- Nsuper*b # Recruits
R


Phi <- js0est$Phi$estimate ## Apparent survival probability
Phi


nYears <- length(R)+1
N <- rep(NA, nYears)
N[1] <- Nsuper*b0 ## Initial abundance
N[2] <- N[1]*Phi + R[1] ## Abundance in year 2
N[3] <- N[2]*Phi + R[2] ## Abundance in year 3
N[4] <- N[3]*Phi + R[3] ## Abundance in year 4
N[5] <- N[4]*Phi + R[4] ## Abundance in year 5
N[6] <- N[5]*Phi + R[5] ## Abundance in year 6
N[7] <- N[6]*Phi + R[6] ## Abundance in year 7
N[8] <- N[7]*Phi + R[7] ## Abundance in year 8
N[9] <- N[8]*Phi + R[8] ## Abundance in year 9
round(N, digits=0)
```

#### **Plot**

```{r, plot2, include=T, warning=F, message=F}

lambda <- N[2:9]/N[1:8]
lam<-round(lambda, digits=16)

Years<-1:8

plot(Years, lam, type="b", xlab="Year", ylab="Lambda")
```